A Responsible Machine Learning Workflow\\\normalsize{With Focus on Interpretable Models, Post-hoc Explanation, and Discrimination Testing}

### 2020-02-05 article-information-2019.tex
bigger figures, checked all external links, confirmed figure an table placement, proofed bibliography and tables/figures

% TODO:
% EDITING:
% - Structure ------------------------- x 
% - External links
% - Internal links
%   - Sections/subsections/appendices - x
%   - References ---------------------- x
%   - Tables/figures
%   - footnotes ----------------------- x
% - Abbreviations --------------------- x
% - Bibliography 
%   - Nick/Bryce citation
% - Notation
% - Tables/figures
% - Grammar and spelling
%   - Hyphenation
%   - "metric" vs. "measure"
% CONTENT:
% - sim data: add error metrics, confirm statements about results ------------------------------------------------------------------------------------------- x
% - sim data xnn: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table --- x
% - mort data: add error metrics, confirm statements about results ----------------------------------------------------------------------------------------- x
% - mort data: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table ------- x
% - make subfigure reference match XNN charts ---------------------------------------------------------------------------------------------------------------- x

@@

Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear. ML models, a common application of AI, can present serious risks. ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.  ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}. This manuscript makes no claim that these interdependent issues of opaqueness, discrimination, privacy harms, and security vulnerabilities in ML have been resolved, even as singular entities, and much less as complex intersectional phenomena. However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available Python tools, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) General Data Protection Regulation (GDPR) Article 22 \cite{ff_interpretability}.}

@@

The same can be said of the selected explanation methods and discrimination tests. Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines. Presented explanation techniques include PD and ICE (see Subsection \ref{ssec:pd_ice}) and Shapley values (see Subsection \ref{ssec:shap}) \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}. PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models. Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}. All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Subsection \ref{ssec:soft}).

@@

\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.  ME is a favored DI metric used by the the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.

@@

\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the proportion of the control class that receives a favorable outcome. Statistically significant AIR values below 0.8 can be considered \textit{prima facia} evidence of discrimination. An additional long-standing measure of DI is SMD. SMD is often used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending. It originates from work on statistical power, and is more formally known as \textit{Cohen’s d}. SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.

@@

Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}. The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.

@@

\begin{table}[H]
	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Arrows indicate the direction of improvement for each measure.}
	\centering
	\begin{tabular}{ccccccccc}
		\toprule
		\textbf{Model} & 
		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ & 
		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
		\midrule
		$g^\text{GBM}$		& 0.757 	& 0.847 	& 0.779 & 0.486 & 0.525 & 0.400 & 0.858 & 0.657 \\
		$g^\text{MGBM}$	& 0.744 	& 0.842 	& 0.771 & 0.502 & 0.504 & 0.407 & 0.864 & 0.625 \\
		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 & 0.525 & 0.398 & 0.858 & 0.657 \\
		$g^\text{XNN}$		& 0.758	& 0.851	& 0.781 & 0.479 & 0.528 & 0.397 & 0.867 & 0.648 \\
		\bottomrule
		\label{tab:sim_acc}
	\end{tabular}
\end{table}

@@

\begin{figure}[H]
	\hskip-20pt\includegraphics[width=18cm]{img/sim_xnn_glob_ridge.png}
	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
	\label{fig:sim_xnn_glob_ridge}
\end{figure}

@@

\begin{table}[htb]
	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data. Arrows indicate the direction of improvement for each measure.}
\centering
\begin{tabular}{ccccccccc}
	\toprule
	\textbf{Model} & 
	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ & 
	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
	\midrule	
$g^\text{GBM}$		& 0.795	& 0.828	& 0.376 & 0.252 & 0.314 & 0.276 & 0.634 & 0.813 \\
$g^\text{MGBM}$	& 0.765	& 0.814	& 0.362 & 0.259 & 0.305 & 0.278 & 0.684 & 0.773 \\
$g^\text{ANN}$		& 0.865	& 0.871	& 0.474 & 0.231 & 0.418 & 0.262 & 0.624 & 0.891 \\
$g^\text{XNN}$		& 0.869	& 0.868	& 0.468 & 0.233 & 0.409 & 0.263 & 0.594 & 0.898 \\
\bottomrule
\label{tab:mort_acc}
\end{tabular}
\end{table}

@@

For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance. Global Shapley feature importance for $g^\text{MGBM}$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}. $g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions. Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the mortgage test data in the $g^\text{MGBM}$ margin space. The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.

\begin{figure}[htb]
\centering
\includegraphics[width=9cm]{img/mort_mgbm_glob.png}
\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data.}
\label{fig:mort_mgbm_glob}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=9cm]{img/mort_xnn_glob.png}
\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
\label{fig:mort_xnn_glob}
\end{figure}

@@

\begin{figure}[htb]
\hskip-20pt\includegraphics[width=18cm]{img/mort_xnn_glob_ridge.png}
\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the mortgage test data.}
\label{fig:mort_xnn_glob_ridge}
\end{figure}

@@

Tables \ref{tab:mort_disc}\textbf{a} and \ref{tab:mort_disc}\textbf{b} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males. As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.  Thus, one might consider FPR ratios as a measure of the class-control disparities. FPR ratios are higher under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$ (2.45 vs. 2.10) in Table \ref{tab:mort_disc}\textbf{b}, but overall FPRs are lower for blacks under $g^{\text{XNN}}$ (0.295 vs. 0.315) in Table \ref{tab:mort_disc}\textbf{a}.  This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.  Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.

>>>>

% TODO:
% EDITING:
% - Structure ------------------------- x 
% - External links -------------------- x
% - Internal links --------------------------- x
%   - Sections/subsections/appendices - x
%   - References ---------------------- x
%   - Tables/figures ----------------- x
%   - footnotes ----------------------- x
% - Abbreviations --------------------- x
% - Bibliography  ---------------------- x 
% - Tables/figures -------------------- x
% - Notation
% - Grammar and spelling
%   - Hyphenation
% CONTENT:
% - sim data: add error metrics, confirm statements about results ------------------------------------------------------------------------------------------- x
% - sim data xnn: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table --- x
% - mort data: add error metrics, confirm statements about results ----------------------------------------------------------------------------------------- x
% - mort data: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table ------- x
% - make subfigure reference match XNN charts ---------------------------------------------------------------------------------------------------------------- x

@@

Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/gx/en/issues/data-and-analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear. ML models, a common application of AI, can present serious risks. ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.  ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}. This manuscript makes no claim that these interdependent issues of opaqueness, discrimination, privacy harms, and security vulnerabilities in ML have been resolved, even as singular entities, and much less as complex intersectional phenomena. However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available Python tools, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) General Data Protection Regulation (GDPR) Article 22 \cite{ff_interpretability}.}

@@

The same can be said of the selected explanation methods and discrimination tests. Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines. Presented explanation techniques include PD and ICE (see Subsection \ref{ssec:pd_ice}) and Shapley values (see Subsection \ref{ssec:shap}) \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}. PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models. Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}. All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Subsection \ref{ssec:soft}).

@@

\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.  ME is a favored DI measure used by the the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.

@@

\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the proportion of the control class that receives a favorable outcome. Statistically significant AIR values below 0.8 can be considered \textit{prima facia} evidence of discrimination. An additional long-standing measure of DI is SMD. SMD is often used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending. It originates from work on statistical power, and is more formally known as \textit{Cohen’s d}. SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this measure to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.

@@

Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}. The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and primarily \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.

@@

\begin{table}[H]
	\caption{Fit measures for $g^\text{GBM}(\mathbf{X})$, $g^\text{MGBM}(\mathbf{X})$, $g^\text{ANN}(\mathbf{X})$, and $g^\text{XNN}(\mathbf{X})$ on the simulated test data. Arrows indicate the direction of improvement for each measure.}
	\centering
	\begin{tabular}{ccccccccc}
		\toprule
		\textbf{Model} & 
		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ & 
		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
		\midrule
		$g^\text{GBM}$		& 0.757 	& 0.847 	& 0.779 & 0.486 & 0.525 & 0.400 & 0.858 & 0.657 \\
		$g^\text{MGBM}$	& 0.744 	& 0.842 	& 0.771 & 0.502 & 0.504 & 0.407 & 0.864 & 0.625 \\
		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 & 0.525 & 0.398 & 0.858 & 0.657 \\
		$g^\text{XNN}$		& 0.758	& 0.851	& 0.781 & 0.479 & 0.528 & 0.397 & 0.867 & 0.648 \\
		\bottomrule
		\label{tab:sim_acc}
	\end{tabular}
\end{table}

@@

\begin{figure}[H]
	\hskip-35pt\includegraphics[width=19cm]{img/sim_xnn_glob_ridge.png}
	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ subnetwork ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the simulated data.}
	\label{fig:sim_xnn_glob_ridge}
\end{figure}

@@

\begin{table}[htb]
	\caption{Fit measures for $g^\text{GBM}(\mathbf{X})$, $g^\text{MGBM}(\mathbf{X})$, $g^\text{ANN}(\mathbf{X})$, and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data. Arrows indicate the direction of improvement for each measure.}
\centering
\begin{tabular}{ccccccccc}
	\toprule
	\textbf{Model} & 
	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ & 
	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
	\midrule	
$g^\text{GBM}$		& 0.795	& 0.828	& 0.376 & 0.252 & 0.314 & 0.276 & 0.634 & 0.813 \\
$g^\text{MGBM}$	& 0.765	& 0.814	& 0.362 & 0.259 & 0.305 & 0.278 & 0.684 & 0.773 \\
$g^\text{ANN}$		& 0.865	& 0.871	& 0.474 & 0.231 & 0.418 & 0.262 & 0.624 & 0.891 \\
$g^\text{XNN}$		& 0.869	& 0.868	& 0.468 & 0.233 & 0.409 & 0.263 & 0.594 & 0.898 \\
\bottomrule
\label{tab:mort_acc}
\end{tabular}
\end{table}

@@

For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance. Global Shapley feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}. $g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions. Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the mortgage test data in the $g^\text{MGBM}$ margin space. The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.

\begin{figure}[htb]
\centering
\captionsetup{width=17cm}
\includegraphics[width=10cm]{img/mort_mgbm_glob.png}
\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data.}
\label{fig:mort_mgbm_glob}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=10cm]{img/mort_xnn_glob.png}
\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
\label{fig:mort_xnn_glob}
\end{figure}

@@

\begin{figure}[htb]
\hskip-35pt\includegraphics[width=19cm]{img/mort_xnn_glob_ridge.png}
\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the mortgage data.}
\label{fig:mort_xnn_glob_ridge}
\end{figure}

@@

Tables \ref{tab:mort_disc}\textbf{a} and \ref{tab:mort_disc}\textbf{b} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males. As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.  Thus, one might consider FPR ratios as a measure of the class-control disparities. FPR ratios are higher under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$ (2.45 vs. 2.10) in Table \ref{tab:mort_disc}\textbf{b}, but overall FPRs are lower for blacks under $g^{\text{XNN}}$ (0.295 vs. 0.315) in Table \ref{tab:mort_disc}\textbf{a}.  This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness measure.  Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
